Fourth Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE4)

### 2017-12-17 WSSSPE4_report/WSSSPE4_paper.tex
incorporated some referee comments

The WSSSPE4 workshop included multiple mechanisms for participation and
encouraged team building around solutions. WSSSPE4 strongly encouraged participation
of early-career scientists, postdoctoral researchers, graduate students,
early-stage researchers, and those from underrepresented groups,
with funds provided to the conference organizers by the National Science
Foundation (NSF), the Gordon and Betty Moore Foundation, the Alfred P.~Sloan Foundation, and the Software
Sustainability Institute (SSI) to support the travel of potential participants
who would not otherwise be able to attend the workshop. These
funds allowed 29 additional people to attend and participate. A subset of the
organizing committee reviewed 44 applications for travel support and
competitively selected the awardees, including 11 students and five early-career
researchers. In addition, the travel award subcommittee tried to increase the
diversity of applicants by forwarding the notice of travel support to
organizations including PyLadies\footnote{\url{http://www.pyladies.com}},
Django Girls\footnote{\url{https://djangogirls.org}},
Women Who Code\footnote{\url{https://www.womenwhocode.com}}, and
Women in HPC\footnote{\url{http://www.womeninhpc.org}}.
%\todo{did we reach out to any others?}
%
WSSSPE4 also included two professional event organizers/facilitators who helped
plan the workshop agenda,
and who engaged with participants during the workshop.

@@

    To make clear what is expected, everyone taking part in WSSSPE events and
    discussions---speakers, helpers, organizers, and participants---is required
    to conform to the following Code of Conduct.

    \begin{itemize}
    \item Communicate appropriately for a professional audience including
    people of many different backgrounds. Sexual language and imagery are not
    appropriate for any event.

    \item Be kind to others. Do not insult or put down other attendees. Be
    mindful of jargon, which can sometimes exclude others from engaging in the
    discussion.

    \item Behave professionally. Remember that harassment and sexist, racist,
    ageist, or exclusionary behavior are not appropriate.
    \end{itemize}
\end{quote}

The CoC was introduced at the beginning of WSSSPE4, with the CoC subcommittee
and a general email address introduced for reporting concerns or incidents, or
asking questions.  There was one concern mentioned to the CoC subcommittee
after the first half day,
regarding how part of the workshop was being run, and we changed the workshop
to address this.

@@

Submissions to WSSSPE4 comprised
19 lightning talks,
4 idea papers,
3 position paper,
5 experience papers,
and
3 demos.
Most of the submissions were accepted, since the goal of WSSSPE is always to
take in as many relevant inputs as possible, and to encourage their authors to
participate in sharing and implementing their ideas.
Specifically,
19 lightning talks,
4 idea papers,
3 position papers,
2 experience papers,
and
3 demos
were accepted, as mentioned in Section~\ref{sec:presentations}. (Two of the submitted lightning talks and one submitted experience paper were rejected, while two more submitted experience papers
were accepted as lightning talks.)
The papers and lightning talks have been published as a volume in the CEUR Workshop Proceedings~\cite{WSSSPE4-proceedings}.

@@

The participants were given time after the presentation to write down their comments.
Seven people volunteered to work on revising the mission and vision statements, incorporating these comments and presenting back to the community.
Based on this feedback, the committee redrafted the mission and vision.
The committee added focus areas to clarify the organization's breadth and to keep the mission and vision simple and long-lasting.
Comments after the close of the meeting were incorporated into the statements.

A final draft was put on GitHub and advertised
via the WSSSPE mailing list, Facebook group, and Twitter.
After two weeks without suggested changes, the final statements, as shown below, were added
to the WSSSPE web page (\url{http://wssspe.researchcomputing.org.uk/about-wssspe/}).

{\bf Mission.}
WSSSPE is an international community-driven organization that promotes sustainable research software by addressing challenges related to the full lifecycle of research software through shared learning and community action.

{\bf Vision.}
We envision a world where research software is accessible, robust, sustained, and recognized as a scholarly research product critical to the advancement of knowledge, learning, and discovery.

{\bf Focus areas.}
WSSSPE promotes sustainable research software by positively impacting:
\begin{itemize}
\item {\bf Principles and Best Practices}. Promoting best practices in sustainable software
\item {\bf Careers}. Developing and supporting career paths in research software development and engineering
\item {\bf Learning}. Engaging in activities to promote peer learning and interaction
\item {\bf Credit}. Ensuring recognition of research software as an intellectual contribution equal to other research products
\end{itemize}

\textbf{Definitions:}
\emph{Sustainable software} has the capacity to endure such that it will continue to
be available in the future, on new platforms, meeting new needs.
The \emph{research software lifecycle} includes:
acquiring and assembling resources (including funding and people) into teams and communities,
developing software,
using software,
recognizing contributions to and of software,
and
maintaining software.

@@

The following four papers belong to the first area: the academic environment.
They address diverse aspects ranging from incentives for quality software to
advocating a professional society for research software to roles and degrees for
research software engineers. For talks with multiple authors, an
\textsuperscript{\textasteriskcentered} indicates the author who presented the talk.

\begin{itemize}
\item Michael Heroux: \emph{Idea paper: Sustainable \& Productive:
Improving Incentives for Quality Software}~\cite{Heroux:2016ws}.

\item Gabrielle Allen: \emph{Idea Paper: Establishing a Professional
Society for Research Software}~\cite{GAllen:2016ws}.

\item Olivier Philippe\textsuperscript{\textasteriskcentered}, Simon Hettrick, and Neil Chue Hong:
\emph{Experience Paper: Preliminary analysis of a survey of UK Research
Software Engineers}~\cite{Philippe:2016ws}.

\item Christopher Gwilliams: \emph{Demo: Using industrial engagement to
create and develop research ties within academia}~\cite{Gwilliams:2016ws}.

\noindent The second main area, concerned with characteristics and needs of research software, includes the following five papers:

\begin{itemize}

\item Anshu Dubey\textsuperscript{\textasteriskcentered} and Katherine Riley: \emph{Experience Paper: Software
Engineering and Community Codes Track in ATPESC}~\cite{Dubey1:2016ws}.

\item Neil Chue Hong: \emph{Position Paper: Why do we need to compare
research software, and how should we do it?}~\cite{ChueHong:2016ws}.

\item Anshu Dubey\textsuperscript{\textasteriskcentered} and Lois Curfman McInnes: \emph{Idea Paper: The Lifecycle
of Software for Scientific Simulation}~\cite{Dubey2:2016ws}.

\item Francisco Queiroz\textsuperscript{\textasteriskcentered} and Rejane Spitz: \emph{Position Paper:
Collaborative Gamification Design for Scientific Software}~\cite{Queiroz:2016ws}.

\item Bruce Childers\textsuperscript{\textasteriskcentered}, Jack Davidson, Wayne Graves, Bernard Rous, and
David Wilkinson: \emph{Position Paper: Active Curation of Artifacts is Changing
the Way Digital Libraries will Operate}~\cite{Childers:2016ws}.

\noindent Three contributions focused on the third main area: elucidating lessons learned
from or visions for use cases of scientific software and communities working
with and\slash or on a software package.

\begin{itemize}

\item Debashis Ganguly, William C.\ Garrison III, David Wilkinson,
Bruce R.\ Childers, Adam J.\ Lee, and Daniel Mosse\textsuperscript{\textasteriskcentered}: \emph{Demo: Composing,
Reproducing, and Sharing Simulations}~\cite{Ganguly:2016ws}.

\item Sameer Shende\textsuperscript{\textasteriskcentered} and Allen Malony: \emph{Demo: Using TAU for
Performance Evaluation of Scientific Software}~\cite{Shende:2016ws}.

\item Janos Sallai\textsuperscript{\textasteriskcentered}, Christopher Iacovella, Christoph Klein, and Tengyu Ma:
\emph{Idea Paper: Development of a Software Framework for Formalizing Forcefield
Atom-Typing for Molecular Simulation}~\cite{Sallai:2016ws}.

@@

\noindent There were 19 lightning talks presented at WSSSPE4, as follows (in order of
presentation).
%
%For talks with multiple authors, an \textsuperscript{\textasteriskcentered} again indicates the author who presented the talk.

\begin{itemize}%[itemsep=1ex]
    \item Shoaib Sufi: \emph{The Software Sustainability Institute Fellowship
    programme; supporting the social side of research software}~\cite{Sufi:2016ws}.

    \item Sandra Gesing\textsuperscript{\textasteriskcentered}, Maytal Dahan, Linda B.~Hayden, Katherine Lawrence,
    Suresh Marru, Marlon Pierce, Nancy Wilkins-Diehr, and Michael Zentner:
    \emph{The Science Gateways Community Institute - Supporting Communities to
    Achieve Sustainability for Their Science Gateways}~\cite{Gesing:2016ws}.

    \item Karthik Ram\textsuperscript{\textasteriskcentered}, Noam Ross,
    and Scott Chamberlain: \emph{A model for
    peer review and onboarding research software}~\cite{Ram:2016ws}.

    \item Frank L\"{o}ffler and Steven Brandt\textsuperscript{\textasteriskcentered}:
    \emph{A vision of computing in 10+ years}~\cite{Loffler:2016ws}.

    \item Daniel S.~Katz, Kyle E.~Niemeyer\textsuperscript{\textasteriskcentered},
    Arfon M.~Smith, and the FORCE11 Software Citation Working Group:
    \emph{Software Citation: Process, Principles, and Implementation}~\cite{Katz:2016ws}.

    \item Ray Idaszak\textsuperscript{\textasteriskcentered},
    David G.~Tarboton, Hong Yi, Michael Stealey,
    Pabitra Dash, Alva Couch, Daniel P.~Ames, Jeffery S.~Horsburgh, Tony Castronova,
    Jon Goodall, Mohamed Morsy, Venkatesh Merwade, Mauriel Ramirez, Tian Gan,
    Drew (Zhiyu) Li, Jeff Sadler, Shawn Crawley, Zhaokun Xue, Lan Zhao, Carol Song,
    Christina Bandaragoda: \emph{HydroShare - A Case Study in Software Engineering
    Best Practices and Culture Change for Developing Sustainable Community
    Software}~\cite{Idaszak:2016ws}.

    \item Lorraine Hwang\textsuperscript{\textasteriskcentered},
    Wolfgang Bangerth, Timo Heister, and Louise Kellogg:
    \emph{ASPECT: Hackathons as an Example of Sustaining an Open Source
    Community}~\cite{Hwang:2016ws}.

    \item Carole Goble: \emph{A Simple Profiling Framework for Software
    User-Producer Reciprocity Review}~\cite{Goble:2016ws}.

    \item Stephan Druskat: \emph{A proposal for the measurement and
    documentation of research software sustainability in interactive metadata
    repositories}~\cite{Druskat:2016ws}.

    \item Mistral Contrastin, Matthew Danish,
    Dominic Orchard\textsuperscript{\textasteriskcentered}, and
    Andrew Rice: \emph{Supporting software sustainability with
    lightweight specifications}~\cite{Contrastin:2016ws}.

    \item Alice Allen, Cecilia Aragon, Christoph Becker, Jeffrey Carver,
    Andrei Chi\c{s}, Benoit Combemale, Mike Croucher, Kevin Crowston, Daniel Garijo,
    Ashish Gehani, Carole Goble\textsuperscript{\textasteriskcentered},
    Robert Haines, Robert Hirschfeld, James Howison,
    Kathryn Huff, Caroline Jay, Daniel S.~Katz, Claude Kirchner, Kateryna Kuksenok,
    Ralf L\"{a}mmel, Oscar Nierstrasz, Matt Turk, Rob van Nieuwpoort, Matthew Vaughn,
    and Jurgen Vinju: \emph{``I solemnly pledge'': A Manifesto for Personal
    Responsibility in the Engineering of Academic Software}~\cite{AAllen:2016ws}.

    \item Neil Chue Hong\textsuperscript{\textasteriskcentered}:
    \textit{Making it easier to understand research software impact}~\cite{ChueHong:2016wsb}.

    \item Willem Robert van Hage\textsuperscript{\textasteriskcentered},
    Jason Maassen and Rob van Nieuwpoort: \textit{Software Impact Measurement at the
    Netherlands eScience Center}~\cite{vanHage:2016ws}.

    \item Gabrielle Allen, Emily Chen\textsuperscript{\textasteriskcentered},
    Ray Idaszak, and Daniel S.\ Katz: \textit{Report on Software Metrics for
    Research Software}~\cite{GAllen:2016wsb}.

    \item Eric L.~Seidel\textsuperscript{\textasteriskcentered}
    and Gabrielle Allen: \textit{Bringing Techniques from Software Engineering
    into Scientific Software}~\cite{Seidel:2016ws}.

    \item Iain Emsley\textsuperscript{\textasteriskcentered} and David
    De Roure: \textit{Sustaining the social: Connecting the lives of Drupal
    community}~\cite{Emsley:2016ws}.

    \item Jack Dongarra, Sven Hammarling, Nicholas J.\ Higham, Samuel
    D.\ Relton\textsuperscript{\textasteriskcentered}, Pedro Valero-Lara,
    and Mawussi Zounon: \textit{Creating a Standardised Set of Batched
    BLAS Routines}~\cite{Dongarra:2016ws}.

    \item Lukas Breitwieser\textsuperscript{\textasteriskcentered},
    Roman Bauer, Alberto Di Meglio, Leonard Johard,
    Marcus Kaiser, Marco Manca, Manuel Mazzara, Fons Rademakers,
    Max Talanov: \textit{The BioDynaMo Project}~\cite{Bauer:2016ws}.

    \item Aseel Aldabjan, Robert Haines, and
    Caroline Jay\textsuperscript{\textasteriskcentered}: \textit{How should we
    measure the relationship between code quality and software
    sustainability?}~\cite{Aldabjan:2016ws}.

@@

Many open source projects for research software document their best practices that
contributed to the sustainability of the software.  Many of these projects also
document software metrics they use to define their research software project's success.
%
This group will take the outputs of the WSSSPE efforts to identify best practices for creation of sustainable software for science/academia, and also the outputs of WSSSPE efforts to identify metrics for sustainable software for science/academia and cross-reference these with current open source research software that successfully uses modern software engineering.  This will allow the group to identify gaps on both sides.  This approach will also allow the group to hypothesize how successful open source projects can be further improved, verify that recommended approaches for software engineering for science/academia are sufficient and valid, and that metrics for software engineering for science/academia are relevant and useful.  The group has a specific objective of getting a good cross-sampling of disparate software including community model codes, community cyberinfrastructure, community analytic tools, etc.

@@

To join this group or obtain more information about it, please send an email to the verifying best practices \& metrics for sustainable research software group: <wssspe4-verify-best-practices@googlegroups.com>.

@@

If interested, visit \url{http://softwaresustainability.org/} or email Neil Chue Hong (<N.ChueHong@software.ac.uk>) and Jean Salac (<jeansalac@virginia.edu>)

@@

To join this project, please visit the GitHub repository at
\url{https://github.com/nunezco2/S2PI} or join the WSSSPE Slack channel
\texttt{\#wg-sci-soft-proto}.

@@

There is more and more academic research being done on topics related to software sustainability, including work on software engineering practices and management of open source projects. However, academic research in general is often siloed for many reasons, and work on topics relevant to software sustainability is no exception. There are many academic studies and projects which may be relevant for practitioners working in this area, and many projects and initiatives by practitioners that may be relevant for social scientists. However, there is a gap between these two in research and practice.
%
This working group is motivated by the goal of building better connections between academic researchers who are studying topics in or relating to software sustainability with practitioners, managers, and administrators who are working in the area of software sustainability.
%
 In any domain, bringing research and practice closer together a mutually beneficial goal, but also has its challenges.
The group aims to give social scientists and practitioners working on scientific software, software sustainability, and open source and a better understanding of each other's work, as well as help them connect and coordinate on specific projects of mutual interest.
%
Furthermore, the group also recognize that academic social science and research software engineering are not monolithic, and there is a need to connect people who care about research-driven best practices inside of these two domains with each other as well.

@@

To join the group, join its mailing list at \url{https://groups.google.com/forum/#!forum/researchsoftwarestudies}

@@

For more information, please contact Emily Chen at <echen35@illinois.edu>.

>>>>

The WSSSPE4 workshop included multiple mechanisms for participation and
encouraged team building around solutions. WSSSPE4 strongly encouraged participation
of early-career scientists, postdoctoral researchers, graduate students,
early-stage researchers, and those from underrepresented groups,
with funds provided to the conference organizers by the National Science
Foundation (NSF), the Gordon and Betty Moore Foundation, the Alfred P.~Sloan Foundation, and the Software
Sustainability Institute (SSI) to support the travel of potential participants
who would not otherwise be able to attend the workshop. These
funds allowed 29 additional people % to attend and participate. A subset of the
% organizing committee reviewed 
out of 44 applicants which 
% ions for travel support and
% competitively selected the awardees, 
included 11 students and five early-career
researchers.
%  In addition, the travel award subcommittee tried to increase the
% diversity of applicants by forwarding the notice of travel support to
% organizations including PyLadies\footnote{\url{http://www.pyladies.com}},
% Django Girls\footnote{\url{https://djangogirls.org}},
% Women Who Code\footnote{\url{https://www.womenwhocode.com}}, and
% Women in HPC\footnote{\url{http://www.womeninhpc.org}}.
%\todo{did we reach out to any others?}
%
WSSSPE4 also included two professional event organizers/facilitators who helped
plan the workshop agenda,
and who engaged with participants during the workshop.

@@

    To make clear what is expected, everyone taking part in WSSSPE events and
    discussions---speakers, helpers, organizers, and participants---is required
    to conform to the Code of Conduct, listed on the community website \cite{}.

%     \begin{itemize}
%     \item Communicate appropriately for a professional audience including
%     people of many different backgrounds. Sexual language and imagery are not
%     appropriate for any event.

%     \item Be kind to others. Do not insult or put down other attendees. Be
%     mindful of jargon, which can sometimes exclude others from engaging in the
%     discussion.

%     \item Behave professionally. Remember that harassment and sexist, racist,
%     ageist, or exclusionary behavior are not appropriate.
%     \end{itemize}
 \end{quote}

The CoC was introduced at the beginning of WSSSPE4, with the CoC subcommittee
and a general email address for reporting concerns or incidents, or
asking questions.  One concern was raised after the first half day, 
% regarding how part of the workshop was being run, 
and we changed the workshop
to address this.

@@

Submissions to WSSSPE4 comprised
19 lightning talks,
4 idea papers,
3 position paper,
5 experience papers,
and
3 demos.
Most of the submissions were accepted, since the goal of WSSSPE is always to
take in as many relevant inputs as possible, and to encourage their authors to
participate in sharing and implementing their ideas.
% Specifically,
% 19 lightning talks,
% 4 idea papers,
% 3 position papers,
% 2 experience papers,
% and
% 3 demos
% were accepted, as mentioned in Section~\ref{sec:presentations}.
(Two of the submitted lightning talks and one submitted experience paper were rejected, while two more submitted experience papers
were accepted as lightning talks.)
The papers and lightning talks have been published as a volume in the CEUR Workshop Proceedings~\cite{WSSSPE4-proceedings}.

@@

The participants were 
%given time after the presentation to write down their  
invited to comment,
% Seven people volunteered to work on 
and 
%revising 
the mission and vision statements, were revised 
%incorporating these comments and presenting back to the community.
based on this feedback.
% the committee redrafted the mission and vision.
The committee added focus areas to clarify the organization's breadth and to keep the mission and vision simple and long-lasting.
Comments after the close of the meeting were also incorporated into the statements.
A final draft was put on GitHub and advertised
via the WSSSPE mailing list, Facebook group, and Twitter.
After two weeks without suggested changes, the final statements, 
%as shown below, 
%were added
to the WSSSPE web page (\url{http://wssspe.researchcomputing.org.uk/about-wssspe/}).

% {\bf Mission.}
% WSSSPE is an international community-driven organization that promotes sustainable research software by addressing challenges related to the full lifecycle of research software through shared learning and community action.

% {\bf Vision.}
% We envision a world where research software is accessible, robust, sustained, and recognized as a scholarly research product critical to the advancement of knowledge, learning, and discovery.

% {\bf Focus areas.}
% WSSSPE promotes sustainable research software by positively impacting:
% \begin{itemize}
% \item {\bf Principles and Best Practices}. Promoting best practices in sustainable software
% \item {\bf Careers}. Developing and supporting career paths in research software development and engineering
% \item {\bf Learning}. Engaging in activities to promote peer learning and interaction
% \item {\bf Credit}. Ensuring recognition of research software as an intellectual contribution equal to other research products
% \end{itemize}

% \textbf{Definitions:}
% \emph{Sustainable software} has the capacity to endure such that it will continue to
% be available in the future, on new platforms, meeting new needs.
% The \emph{research software lifecycle} includes:
% acquiring and assembling resources (including funding and people) into teams and communities,
% developing software,
% using software,
% recognizing contributions to and of software,
% and
% maintaining software.

@@

% The following four papers belong to the first area: the academic environment.
%For talks with multiple authors, an
% \textsuperscript{\textasteriskcentered} indicates the author who presented the talk.
Four papers, 
% \begin{itemize}
% \item Michael Heroux: \emph{Idea paper: Sustainable \& Productive:
% Improving Incentives for Quality Software}
~\cite{Heroux:2016ws}, 
% \item Gabrielle Allen: \emph{Idea Paper: Establishing a Professional
% Society for Research Software}
~\cite{GAllen:2016ws}, 
% \item Olivier Philippe\textsuperscript{\textasteriskcentered}, Simon Hettrick, and Neil Chue Hong:
% \emph{Experience Paper: Preliminary analysis of a survey of UK Research
% Software Engineers}
~\cite{Philippe:2016ws}, and 
% \item Christopher Gwilliams: \emph{Demo: Using industrial engagement to
% create and develop research ties within academia}
~\cite{Gwilliams:2016ws} belong to the first area: the academic environment.
They address diverse aspects ranging from incentives for quality software to
advocating a professional society for research software to roles and degrees for
research software engineers.
%\end{itemize}
The second main area, concerned with characteristics and needs of
research software, includes five papers:
~\cite{Dubey1:2016ws},  ~\cite{ChueHong:2016ws},
~\cite{Dubey2:2016ws}, ~\cite{Queiroz:2016ws}, and ~\cite{Childers:2016ws}.
% \begin{itemize}
% \item Anshu Dubey\textsuperscript{\textasteriskcentered} and Katherine Riley: \emph{Experience Paper: Software
% Engineering and Community Codes Track in ATPESC}
% \item Neil Chue Hong: \emph{Position Paper: Why do we need to compare
% research software, and how should we do it?}
% \item Anshu Dubey\textsuperscript{\textasteriskcentered} and Lois Curfman McInnes: \emph{Idea Paper: The Lifecycle
% of Software for Scientific Simulation}
% \item Francisco Queiroz\textsuperscript{\textasteriskcentered} and Rejane Spitz: \emph{Position Paper:
% Collaborative Gamification Design for Scientific Software}
% \item Bruce Childers\textsuperscript{\textasteriskcentered}, Jack Davidson, Wayne Graves, Bernard Rous, and
% David Wilkinson: \emph{Position Paper: Active Curation of Artifacts is Changing
% the Way Digital Libraries will Operate}
% \end{itemize}
%\noindent 
Three contributions, ~\cite{Ganguly:2016ws}, ~\cite{Shende:2016ws},
and ~\cite{Sallai:2016ws} , focused on the third main area: elucidating lessons learned
from or visions for use cases of scientific software and communities working
with and\slash or on a software package. There were 19 lightning talks presented at WSSSPE4, (in order of
presentation): ~\cite{Sufi:2016ws}, ~\cite{Gesing:2016ws},
~\cite{Ram:2016ws}, ~\cite{Loffler:2016ws}, ~\cite{Katz:2016ws},
~\cite{Idaszak:2016ws}, ~\cite{Hwang:2016ws}, ~\cite{Goble:2016ws},
~\cite{Druskat:2016ws}, ~\cite{Contrastin:2016ws},
~\cite{AAllen:2016ws}, ~\cite{ChueHong:2016wsb},
~\cite{vanHage:2016ws}, ~\cite{GAllen:2016wsb}, ~\cite{Seidel:2016ws}, 
  ~\cite{Emsley:2016ws}, ~\cite{Dongarra:2016ws},
  ~\cite{Bauer:2016ws},  and ~\cite{Aldabjan:2016ws}.

% \begin{itemize}
% \item Debashis Ganguly, William C.\ Garrison III, David Wilkinson,
% Bruce R.\ Childers, Adam J.\ Lee, and Daniel Mosse\textsuperscript{\textasteriskcentered}: \emph{Demo: Composing,
% Reproducing, and Sharing Simulations}
% \item Sameer Shende\textsuperscript{\textasteriskcentered} and Allen Malony: \emph{Demo: Using TAU for
% Performance Evaluation of Scientific Software}
% \item Janos Sallai\textsuperscript{\textasteriskcentered}, Christopher Iacovella, Christoph Klein, and Tengyu Ma:
% \emph{Idea Paper: Development of a Software Framework for Formalizing Forcefield
% Atom-Typing for Molecular Simulation}
% \end{itemize}

@@

%
%For talks with multiple authors, an \textsuperscript{\textasteriskcentered} again indicates the author who presented the talk.

% \begin{itemize}%[itemsep=1ex]
%     \item Shoaib Sufi: \emph{The Software Sustainability Institute Fellowship
%     programme; supporting the social side of research software}

%     \item Sandra Gesing\textsuperscript{\textasteriskcentered}, Maytal Dahan, Linda B.~Hayden, Katherine Lawrence,
%     Suresh Marru, Marlon Pierce, Nancy Wilkins-Diehr, and Michael Zentner:
%     \emph{The Science Gateways Community Institute - Supporting Communities to
%     Achieve Sustainability for Their Science Gateways}

%     \item Karthik Ram\textsuperscript{\textasteriskcentered}, Noam Ross,
%     and Scott Chamberlain: \emph{A model for
%     peer review and onboarding research software}
%     \item Frank L\"{o}ffler and Steven Brandt\textsuperscript{\textasteriskcentered}:
%     \emph{A vision of computing in 10+ years}
%     \item Daniel S.~Katz, Kyle E.~Niemeyer\textsuperscript{\textasteriskcentered},
%     Arfon M.~Smith, and the FORCE11 Software Citation Working Group:
%     \emph{Software Citation: Process, Principles, and
%       Implementation}

%     \item Ray Idaszak\textsuperscript{\textasteriskcentered},
%     David G.~Tarboton, Hong Yi, Michael Stealey,
%     Pabitra Dash, Alva Couch, Daniel P.~Ames, Jeffery S.~Horsburgh, Tony Castronova,
%     Jon Goodall, Mohamed Morsy, Venkatesh Merwade, Mauriel Ramirez, Tian Gan,
%     Drew (Zhiyu) Li, Jeff Sadler, Shawn Crawley, Zhaokun Xue, Lan Zhao, Carol Song,
%     Christina Bandaragoda: \emph{HydroShare - A Case Study in Software Engineering
%     Best Practices and Culture Change for Developing Sustainable Community
%     Software}

%     \item Lorraine Hwang\textsuperscript{\textasteriskcentered},
%     Wolfgang Bangerth, Timo Heister, and Louise Kellogg:
%     \emph{ASPECT: Hackathons as an Example of Sustaining an Open Source
%     Community}

%     \item Carole Goble: \emph{A Simple Profiling Framework for Software
%     User-Producer Reciprocity Review}
%     \item Stephan Druskat: \emph{A proposal for the measurement and
%     documentation of research software sustainability in interactive metadata
%     repositories}

%     \item Mistral Contrastin, Matthew Danish,
%     Dominic Orchard\textsuperscript{\textasteriskcentered}, and
%     Andrew Rice: \emph{Supporting software sustainability with
%     lightweight specifications}

%     \item Alice Allen, Cecilia Aragon, Christoph Becker, Jeffrey Carver,
%     Andrei Chi\c{s}, Benoit Combemale, Mike Croucher, Kevin Crowston, Daniel Garijo,
%     Ashish Gehani, Carole Goble\textsuperscript{\textasteriskcentered},
%     Robert Haines, Robert Hirschfeld, James Howison,
%     Kathryn Huff, Caroline Jay, Daniel S.~Katz, Claude Kirchner, Kateryna Kuksenok,
%     Ralf L\"{a}mmel, Oscar Nierstrasz, Matt Turk, Rob van Nieuwpoort, Matthew Vaughn,
%     and Jurgen Vinju: \emph{``I solemnly pledge'': A Manifesto for Personal
%     Responsibility in the Engineering of Academic
%     Software}

%     \item Neil Chue Hong\textsuperscript{\textasteriskcentered}:
%     \textit{Making it easier to understand research software
%       impact}

%     \item Willem Robert van Hage\textsuperscript{\textasteriskcentered},
%     Jason Maassen and Rob van Nieuwpoort: \textit{Software Impact Measurement at the
%     Netherlands eScience Center}

%     \item Gabrielle Allen, Emily Chen\textsuperscript{\textasteriskcentered},
%     Ray Idaszak, and Daniel S.\ Katz: \textit{Report on Software Metrics for
%     Research Software}

%     \item Eric L.~Seidel\textsuperscript{\textasteriskcentered}
%     and Gabrielle Allen: \textit{Bringing Techniques from Software Engineering
%     into Scientific Software}

%     \item Iain Emsley\textsuperscript{\textasteriskcentered} and David
%     De Roure: \textit{Sustaining the social: Connecting the lives of Drupal
%     community}

%     \item Jack Dongarra, Sven Hammarling, Nicholas J.\ Higham, Samuel
%     D.\ Relton\textsuperscript{\textasteriskcentered}, Pedro Valero-Lara,
%     and Mawussi Zounon: \textit{Creating a Standardised Set of Batched
%     BLAS Routines}

%     \item Lukas Breitwieser\textsuperscript{\textasteriskcentered},
%     Roman Bauer, Alberto Di Meglio, Leonard Johard,
%     Marcus Kaiser, Marco Manca, Manuel Mazzara, Fons Rademakers,
%     Max Talanov: \textit{The BioDynaMo Project}

%     \item Aseel Aldabjan, Robert Haines, and
%     Caroline Jay\textsuperscript{\textasteriskcentered}: \textit{How should we
%     measure the relationship between code quality and software
%     sustainability?}

% \end{itemize}

@@

Many open source projects for research software document their best practices that
contribute to the sustainability of the software.  Many of these projects also
document software metrics they use to define their research software project's success.
%
This group will take the outputs of the WSSSPE efforts to identify best practices for creation of sustainable software for science/academia, and also the outputs of WSSSPE efforts to identify metrics for sustainable software for science/academia and cross-reference these with current open source research software that successfully uses modern software engineering.  This will allow the group to identify gaps on both sides.  This approach will also allow the group to hypothesize how successful open source projects can be further improved, verify that recommended approaches for software engineering for science/academia are sufficient and valid, and that metrics for software engineering for science/academia are relevant and useful.  The group has a specific objective of getting a good cross-sampling of disparate software including community model codes, community cyberinfrastructure, community analytic tools, etc.

@@

Anyone can join this group or obtain more information about it, by sending an email to the verifying best practices \& metrics for sustainable research software group: <wssspe4-verify-best-practices@googlegroups.com>.

@@

If interested, one can visit \url{http://softwaresustainability.org/} or email Neil Chue Hong (<N.ChueHong@software.ac.uk>) and Jean Salac (<jeansalac@virginia.edu>)

@@

To join this project, one can visit the GitHub repository at
\url{https://github.com/nunezco2/S2PI} or join the WSSSPE Slack channel
\texttt{\#wg-sci-soft-proto}.

@@

There is more and more academic research being done on topics related to software sustainability, including work on software engineering practices and management of open source projects. However, academic research in general is often siloed for many reasons, and work on topics relevant to software sustainability is no exception. There are many academic studies and projects which may be relevant for practitioners working in this area, and many projects and initiatives by practitioners that may be relevant for social scientists. However, there is a gap between these two in research and practice.
%
This working group is motivated by the goal of building better connections between academic researchers who are studying topics in or relating to software sustainability with practitioners, managers, and administrators who are working in the area of software sustainability.
%
 In any domain, bringing research and practice closer together a mutually beneficial goal, but also has its challenges.
The group aims to give social scientists and practitioners working on scientific software, software sustainability, and open source and a better understanding of each other's work, as well as help them connect and coordinate on specific projects of mutual interest.
%
Furthermore, the group also recognizes that academic social science and research software engineering are not monolithic, and there is a need to connect people who care about research-driven best practices inside of these two domains with each other as well.

@@

To join the group, one can join its mailing list at \url{https://groups.google.com/forum/#!forum/researchsoftwarestudies}

@@

For more information, one can contact Emily Chen at <echen35@illinois.edu>.
